Can we use masked priming in an online setting?

### 2021-05-05 Lab meeting presentation April 2021 Results.Rmd
Added general discussion, edited introduction

- `r exp1_all_participants$PROLIFIC_PID %>% unique %>% length` Prolific participants, with `r exp1_accuracy_by_participant %>% nrow` with `r exp1_actual_participants$PROLIFIC_PID %>% unique %>% length` completing the experiment.
- Participants (`r exp1_male_participants_nr` male, `r exp1_female_participants_nr` female) were aged from `r exp1_participant_ages %>% min` to `r exp1_participant_ages %>% max` (mean age: `r exp1_participant_ages %>% mean %>% round(2)`). Of the participants,  identified as male and `r exp1_female_participants_nr` identified as female. The experiment was only shown to participants who indicated that English was their first language in the Prolific screening questions. Based on their IP addresses, `r exp1_participant_countries["United Kingdom"] %>% ifelse(. < 11, numbers_to_words(.), .)` participants were based in the UK, `r exp1_participant_countries["United States of America"] %>% ifelse(. < 11, numbers_to_words(.), .)` were based in the US, `r exp1_participant_countries["Canada"] %>% ifelse(. < 11, numbers_to_words(.), .)` participants were based in Canada, and `r exp1_participant_countries["Ireland"] %>% ifelse(. < 11, numbers_to_words(.), .)` participants were based in Ireland. Two participants could not be localized in this way. Participants who completed the experiment were paid £1.25 for their participation (corresponding to £5/hour). All participants were naïve to the purpose of the experiment. Participants could use a desktop or laptop computer or a mobile device. Because of a technical display issue with PsychoJS and the Safari browser, participants who tried to access the experiment using that browser, including all participants on iOS devices, were advised to change browser or device and restart the experiment.

>>>>

- `r exp1_all_participants$PROLIFIC_PID %>% unique %>% length` Prolific participants, with `r exp1_accuracy_by_participant %>% nrow` with `r exp1_actual_participants$PROLIFIC_PID %>% unique %>% length` completing the experiment.
- Participants (`r exp1_male_participants_nr` male, `r exp1_female_participants_nr` female) were aged from `r exp1_participant_ages %>% min` to `r exp1_participant_ages %>% max` (mean age: `r exp1_participant_ages %>% mean %>% round(2)`)s. Of the participants,  identified as male and `r exp1_female_participants_nr` identified as female. The experiment was only shown to participants who indicated that English was their first language in the Prolific screening questions. Based on their IP addresses, `r exp1_participant_countries["United Kingdom"] %>% ifelse(. < 11, numbers_to_words(.), .)` participants were based in the UK, `r exp1_participant_countries["United States of America"] %>% ifelse(. < 11, numbers_to_words(.), .)` were based in the US, `r exp1_participant_countries["Canada"] %>% ifelse(. < 11, numbers_to_words(.), .)` participants were based in Canada, and `r exp1_participant_countries["Ireland"] %>% ifelse(. < 11, numbers_to_words(.), .)` participants were based in Ireland. Two participants could not be localized in this way. Participants who completed the experiment were paid £1.25 for their participation (corresponding to £5/hour). All participants were naïve to the purpose of the experiment. Participants could use a desktop or laptop computer or a mobile device. Because of a technical display issue with PsychoJS and the Safari browser, participants who tried to access the experiment using that browser, including all participants on iOS devices, were advised to change browser or device and restart the experiment.

### 2021-05-05 Manuscript_intro_method_results.Rmd
Added general discussion, edited introduction

In the present study, we were interested in whether we could replicate, and extend, a key phenomenon in laboratory masked priming lexical decision using an online setup: masked identity priming reflects a savings effect. As first suggested by @forster1998pros, for a masked identity prime, "the lexical entry is already in the process of being opened, and hence the evaluation of this entry begins sooner", whereas for an unrelated prime, "the entry for the target word would be closed down (since it fails to match the prime), and no savings would occur" (p. 213). Thus, according to the savings account, a target word like *DOCTOR* would enjoy an encoding advantage when preceded by an identity prime such as *doctor* than when preceded by an unrelated prime such as *pencil* (i.e., a head-start). One implication of such benefit is that the RT distributions of the unrelated and identity pairs should reflect a shift rather than a change in shape. Furthermore, this shift should be approximately similar in magnitude to the prime-target stimulus-onset asynchrony. Empirical evidence supporting this view has been obtained in several papers not only with skilled adults but also with developing readers [e.g., @gomezDiffusionModelAccount2013; @Gomez_2020; @Taikh_2020; @Yang_2021]. Moreover, this pattern of results fits very well with the diffusion model [@ratcliff2004]. This model proposes that, when making a two-choice decision, the resultant RT can be explained as the sum of non-decision parameters, which are the encoding time and response execution ($T_{er}$) and decision parameters, which refer to the process of accumulation of information until a decision criteria is reached. Importantly, in the decision process, the information gathered from the stimulus can vary in noise, depending on its quality, which modifies the rate at which information is accumulated (i.e., the *drift rate*). With regards to RTs from masked priming tasks, @gomezDiffusionModelAccount2013 found that the difference between identity and unrelated conditions could be accounted by a change in the $T_{er}$ parameter, while there were no differences across conditions in the parameter that corresponds to the quality of evidence gathered (i.e., drift rate)--note that changes in drift rate would necessarily produce a skewer RT distribution in the slower, unrelated condition.

@@

In the present paper, we took advantage of the above marker to examine whether online masked priming studies follow the same pattern as in-lab masked priming studies. Specifically, we manipulated prime exposure duration in identity vs. unrelated primes: 33.3 vs. 50 ms in Experiment 1, and 16.6 vs. 33.3 ms in Experiment 2---note that targets were presented immediately after the primes (i.e., prime exposure duration was equal to the prime-target SOA). The rationale of Experiment 1 is that if the actual exposure duration of the primes is the nominal exposure duration, then we would expect the typical shift between the identity and unrelated response time distributions, which according to the savings hypothesis [@forster1998pros] would be greater for 50-ms than for 33.3-ms exposure duration (i.e., the head-start would be greater for 50-ms identity primes than for 33.3-ms identity primes). This outcome would indicate that the on-line masked priming studies reproduce a characteristic signature of laboratory masked priming studies. Alternatively, if the actual exposure duration is greater than the nominal one (e.g., approximately 20-ms longer on average) then, the 50 ms primes could be consciously perceived. If this were the case, the effect of the prime could affect not only the encoding but also core decision processes (i.e., drift rate parameter), which would be reflected as a stronger priming effect in the upper quantiles (i.e., the two RT distributions would have a different shape). In this scenario, one should be very cautious when running online masked priming experiments---at least with the settings currently available.

To further constraint the research questions, Experiment 2 was parallel to Experiment 1, except for the use of a very short prime exposure duration (i.e., 16.6ms). Similar prime exposure durations have shown rather weak masked priming lexical decision experiments in a laboratory setting: smaller than 5 ms for prime durations of 14 ms [@ziegler2000] and smaller than 9 ms for a prime duration of 20 ms [@Tzur_2007]---in the @Tzur_2007 experiment, this difference increased to 16 ms when using a very high level of contrast in the computer screen. Thus, if prime exposure durations in online masked priming correspond to the nominal ones, we would expect a sizeable priming effect at the 33.3 but only a residual priming effect at the 16.6 ms prime exposure duration. Alternatively, if we obtain a sizeable effect at the 16.6 ms prime exposure duration (e.g., above 20-25 ms), this would indicate, again, that we should be wary of using the masked priming technique in online experiments. Keep in mind that we are using a software that has good control over the exposure duration (\color{red} CITE).

@@

```{r exp1-blmm-table, echo=FALSE, as.is = TRUE}
# We fit the models in a separate file to make sure that knitting the manuscript doesn't take too long
load("blmm_exp1_rt_new.RData")
# model name is blmm_exp1_rt

@@

blmm_exp1_summary_rt <- summary(blmm_exp1_rt)

@@

@@

apa_table(
    blmm_exp1_rt_fixed_for_print,
    placement = "H",
    align = c("c", "c", "c", "c", "c"),
    caption = "(ref:exp1-blmm-table-caption)",
    note = NULL,
    small = TRUE,
    digits = 3,
    escape = FALSE
    )

@@

load("blmm_exp1_acc_new.RData")

@@

The results from Experiment 1 show that we were able to replicate masked priming effects using an online experiment. The size of the effect is similar to that observed in previous studies. We also saw a clear difference between the 33 ms prime duration and the 50 ms prime duration, indicating that the experiment can reliably implement timing differences of up to one frame across a variety of participant devices. Of course, just because there was a difference between the conditions, this does not necessarily mean that the timings in the two prime duration conditions actually corresponded to the display durations set in the experiment script, just that they were different. In order to further explore this, we perfomed a second experiment in which we set the prime to be displayed for an even shorter duration. As described above, a 16 ms prime duration should yield qualitatively different effects from the 33 and 50 ms durations used in Experiment 1. If it does not, this casts doubt on the timing accuracy in online experiments.

@@

```{r load_exp2_brms_data, echo=FALSE, as.is = TRUE}
# We fit the models in a separate file to make sure that knitting the manuscript doesn't take too long
load("blmm_exp2_rt_new.RData")
# model name is blmm_exp1_rt

@@

blmm_exp2_summary_rt <- summary(blmm_exp2_rt)

@@

apa_table(
    blmm_exp2_rt_fixed_for_print,
    placement = "H",
    align = c("c", "c", "c", "c", "c"),
    caption = "(ref:exp2-blmm-rt-table-caption)",
    note = NULL,
    small = TRUE,
    digits = 3,
    escape = FALSE
    )

@@

load("blmm_exp2_acc_new.RData")

@@

In Experiment 2, we observed the pattern we expected from previous research: a robust priming effect in response time and accuracy in the 33 ms condition, but a very weak effect in the 16 ms condition. This suggests that the timing in our online experiments was likely to be quite close to the timing set in the experiment script.

@@

In this study, we

>>>>

In the present study, we were interested in whether we could replicate, and extend, a key phenomenon in laboratory masked priming lexical decision using an online setup: masked identity priming reflects a savings effect. As first suggested by @forster1998pros, for a masked identity prime, "the lexical entry is already in the process of being opened, and hence the evaluation of this entry begins sooner", whereas for an unrelated prime, "the entry for the target word would be closed down (since it fails to match the prime), and no savings would occur" (p. 213). Thus, according to the savings account, a target word like *DOCTOR* would enjoy an encoding advantage when preceded by an identity prime such as *doctor* than when preceded by an unrelated prime such as *pencil* (i.e., a head-start). One implication of such benefit is that the RT distributions of the unrelated and identity pairs should reflect a shift rather than a change in shape. Furthermore, this shift should be approximately similar in magnitude to the prime-target stimulus-onset asynchrony (SOA). Empirical evidence supporting this view has been obtained in several papers not only with skilled adults but also with developing readers [e.g., @gomezDiffusionModelAccount2013; @Gomez_2020; @Taikh_2020; @Yang_2021]. Moreover, this pattern of results fits very well with the diffusion model [@ratcliff2004]. This model proposes that, when making a two-choice decision, the resultant RT can be explained as the sum of non-decision parameters, which are the encoding time and response execution ($T_{er}$) and decision parameters, which refer to the process of accumulation of information until a decision criteria is reached. Importantly, in the decision process, the information gathered from the stimulus can vary in noise, depending on its quality, which modifies the rate at which information is accumulated (i.e., the *drift rate*). With regards to RTs from masked priming tasks, @gomezDiffusionModelAccount2013 found that the difference between identity and unrelated conditions could be accounted by a change in the $T_{er}$ parameter, while there were no differences across conditions in the parameter that corresponds to the quality of evidence gathered (i.e., drift rate)--note that changes in drift rate would necessarily produce a skewer RT distribution in the slower, unrelated condition.

@@

In the present paper, we took advantage of the above marker to examine whether online masked priming studies follow the same pattern as in-lab masked priming studies. Specifically, we manipulated prime exposure duration in identity vs. unrelated primes: 33.3 vs. 50 ms in Experiment 1, and 16.6 vs. 33.3 ms in Experiment 2---note that targets were presented immediately after the primes (i.e., prime exposure duration was equal to the prime-target SOA). The rationale of Experiment 1 is that if the actual exposure duration of the primes is the nominal exposure duration, then we would expect the typical shift between the identity and unrelated response time distributions, which according to the savings hypothesis [@forster1998pros] would be greater for 50 ms than for 33.3 ms exposure duration (i.e., the head-start would be greater for 50 ms identity primes than for 33.3 ms identity primes). This outcome would indicate that the on-line masked priming studies reproduce a characteristic signature of laboratory masked priming studies. Alternatively, if the online presentation conditions lead to a greater actual exposure duration on the participant's device compared to that specified in our experiment, (e.g., prime durations are approximately 20 ms longer on average than programmed) then the 50 ms primes may no longer be masked, but may rather be consciously perceived. If this is the case, the prime could affect not only the encoding, but also core decision processes (i.e., the drift rate), which would be reflected as a stronger priming effect in the higher quantiles of the distribution (i.e., the two RT distributions would have a different shape). In this scenario, one should be very cautious when running online masked priming experiments---at least with the typical software and hardware currently available.

To further constrain the research questions, Experiment 2 was designed to be analogous to Experiment 1, except for the use of a very short prime exposure duration (i.e., 16.6ms). Similar prime exposure durations have shown rather weak masked priming lexical decision experiments in a laboratory setting: smaller than 5 ms for prime durations of 14 ms [@ziegler2000] and smaller than 9 ms for a prime duration of 20 ms [@Tzur_2007]---in the @Tzur_2007 experiment, this difference increased to 16 ms when using a very high level of contrast in the computer screen. Thus, if the size of the priming effect is roughly similar to the prime exposure duration, we would expect a much larger priming effect at the 33.3 ms prime exposure duration than at the 16.6 ms prime exposure duration. If we do observe a large effect at the 16.6 ms prime exposure duration (e.g., above 20-25 ms), this would suggest, again, that there is a qualitative difference between using the masked priming technique in the laboratory and in online experiments. Keep in mind that we are using a software that has good control over the exposure duration (\color{red} CITE).

@@

```{r exp1-blmm-table, echo=FALSE, as.is = TRUE}
# We fit the models in a separate file to make sure that knitting the manuscript doesn't take too long
load("blmm_exp1_rt_dist_full_ranef_5000.RData")
# model name is blmm_exp1_rt

@@

blmm_exp1_summary_rt <- summary(blmm_exp1_rt_dist)

@@

parameter_names <- c(
    "Intercept ($\\mu$)",
    "Intercept ($\\beta$)",
    "Condition ($\\mu$)",
    "Prime Duration ($\\mu$)",
    "Condition:Prime Duration ($\\mu$)",
    "Condition ($\\beta$)",
    "Prime Duration ($\\beta$)",
    "Condition:Prime Duration ($\\beta$)"
    )

@@

apa_table(
    blmm_exp1_rt_fixed_for_print,
    placement = "H",
    align = c("c", "c", "c", "c", "c"),
    caption = "(ref:exp1-blmm-table-caption)",
    note = "$\\beta$ is the scale parameter (the inverse of the rate parameter $\\lambda$) of the ex-Gaussian distribution.",
    small = TRUE,
    digits = 3,
    escape = FALSE
    )

@@

load("blmm_exp1_acc_new_10000.RData")

@@

The results from Experiment 1 show that we were able to replicate masked priming effects using an online experiment. The size of the effect is similar to that observed in previous studies. We also saw a clear difference between the 33 ms prime duration and the 50 ms prime duration, indicating that the experiment can reliably implement timing differences of up to one frame across a variety of participant devices. We did not observe an effect of either the priming condition or the prime exposure duration on the shape parameter $\beta$ of the exponential distribution, suggesting, according to the savings hypothesis by @forster1998pros, that only encoding processes were affected.

Of course, just because there was a difference between the conditions, this does not necessarily mean that the timings in the two prime duration conditions actually corresponded to the display durations set in the experiment script, just that they were different. In order to further explore this, we performed a second experiment in which we set the prime to be displayed for an even shorter duration. As described above, a 16.6 ms prime duration should yield qualitatively different effects from the 33.3 and 50 ms durations used in Experiment 1. If it does not, this casts doubt on the timing accuracy in online experiments.

@@

```{r load_exp2_brms_data, echo=FALSE, as.is = TRUE}
# We fit the models in a separate file to make sure that knitting the manuscript doesn't take too long
load("blmm_exp2_rt_dist.RData")
# model name is blmm_exp2_rt

@@

blmm_exp2_summary_rt <- summary(blmm_exp2_rt_dist)

@@

apa_table(
    blmm_exp2_rt_fixed_for_print,
    placement = "H",
    align = c("c", "c", "c", "c", "c"),
    caption = "(ref:exp2-blmm-rt-table-caption)",
    note = "$\\beta$ is the scale parameter (the inverse of the rate parameter $\\lambda$) of the ex-Gaussian distribution.",
    small = TRUE,
    digits = 3,
    escape = FALSE
    )

@@

load("blmm_exp2_acc_new_10000.RData")

@@

In Experiment 2, we observed the pattern we expected from previous research: In the the mean $\mu$ of the Gaussian distribution, we saw a robust priming effect in response time and accuracy in the 33 ms condition, but a very weak effect in the 16 ms condition. This suggests that the timing in our online experiments was likely to be quite close to the timing set in the experiment script. Again, we did not observe an effect of either the priming condition or the prime exposure duration on the shape parameter $\beta$ of the exponential distribution.

@@

In this study, we set out to test whether using an online, browser-based experiment software, we could replicate the masked priming effect both qualitatively and quantitatively, i.e., in terms of effect sizes. Our results show that this is the case: not only did we observe results comparable to the lab-based studies in the literature, we also observed the effect sizes predicted by Forster's [-@forster1998pros] savings hypothesis: Our data show a shift in the mean of the Gaussian distribution, but no change in the shape parameter $\beta$ of the exponential distribution, suggesting that our priming manipulations affected --- as intended --- encoding processes, but not conscious decision making processes.

We also found that the size of the priming effect was directly influenced by the prime exposure duration. This indicates that the experimental software was able to control the display timing of the prime accurately. Accurate display timings are expected in a laboratory-based experiment, where the equipment is known and can be measured, but are much less certain in a situation where the experiment runs on a participant's own device, which could be any of a wide variety of consumer devices including Windows PCs, Macs, tablets, and mobile phones sold in the last decade. The fact that we can observe results that very closely resemble lab results demonstrates the sophistication in modern browsers' Javascript performance, including browsers on mobile devices, as well as the quality of the Javascript implementation of Psychopy.

Our results open the door to a wider use of online experiments in cognitive research, especially in reaction-time sensitive fields like word recognition. Not only does this enable researchers to continue collecting data in times of social distancing, but it also makes it possible to collect data from a larger population than previously possible. For example, many people in developing countries may not have computers at home, but they do have smartphones. Using a Javascript-based experiment software, anyone with a smartphone can be an experiment participant. Deploying experiments online also means that data can be collected very quickly and efficiently, allowing research to progress more rapidly.

In conclusion, our results give us confidence that high-quality behavioral data can be collected online using Javascript-based experiment platforms. We hope that future research takes advantage of these new methods in order to make research faster, more inclusive, and more efficient.
